The low-frequency Murchison radio environment:\\interference analysis and mitigation

### 2014-10-14 articles/2014/MWA-Radio-Environment/mwa-radio-environment.tex
Work

\begin{table*}
\caption{List of observations used in the analyses. The RFI column contains the fraction of visibilities that the initial \textsc{aoflagger} run has classified as RFI (which includes false positives). The DVB column displays the fraction of visibilities that were unusable because of the presence of DVB signals, and which the flagger does not flag. Not all observations cover the DVB frequencies. }\label{tbl:obs-list}
\begin{center}%
\begin{tabular}{|l|l|r|r|r|c|}
\hline
\textbf{Project} &\textbf{Date} & \textbf{Frequencies (MHz)} & \textbf{Duration} & RFI & DVB \\
\hline
EoR high& 2013-08-23 & 167.0--197.7              & 3h & 0.53\% & 0\% \\
GLEAM & 2013-08-25 & 72.3--133.8, 138.9--230.8 & 7h & 0.94\% & 0\% \\
EoR low& 2013-08-26 & 138.9--169.6              & 6h & 0.81\% & ---\\
GLEAM & 2013-11-05 & 72.3--133.8, 138.9--230.8 & 7h & 1.27\% & 0\% \\
GLEAM & 2013-11-25 & 72.3--133.8, 138.9--230.8 & 7h & 0.69\% & 0\% \\
EoR high& 2014-02-05 & 167.0--197.7              & 6h & 0.54\% & 0\%\\
GLEAM & 2014-03-16 & 72.3--133.8, 138.9--230.8 & 7h & 0.79\% & 0\% \\
GLEAM & 2014-03-17 & 72.3--133.8, 138.9--230.8 & 7h & 1.64\% & 30\% \\
EoR high& 2014-04-10 & 167.0--197.7              & 6h & 0.68\% & 0\%\\
GLEAM & 2014-06-18 & 72.3--133.8, 138.9--230.8 & 7h & 0.98\% & 0\% \\
Orbcomm RFI test&2014-08-27& 131.2--161.9     & 2$\times$8 m& 1.85\% & --- \\
\hline
\multicolumn{3}{|r|}{\textbf{Total}} & \textbf{63h} & \textbf{0.96\%} & \textbf{3\%}\\
\hline
\multicolumn{4}{|r|}{\textbf{With uniform channel weighting}} & \textbf{1.13\%} & \\
\hline
\end{tabular}
\end{center}
\end{table*}

@@

The fifth column of Table~\ref{tbl:obs-list} lists for each of the observation nights the ratio of samples that were classified as RFI by the \textsc{aoflagger}. This includes false detections, which are estimated in \citet{lofar-radio-environment} to account for approximately 0.4\% detections. Compared to the EoR observations, the GLEAM observations have a higher average of 1.05\% of RFI. This is mostly caused by the FM bands, which are only observed in GLEAM observations. The EoR low-band observation shows 0.81\% RFI, while the EoR high-band observations have an average of 0.58\%. The total RFI occupancy in all observations is 0.96\%. Weighting the occupancy in each channel by the inverse time that it is observed results in an occupancy of 1.13\%, which is the global expected RFI ratio in the 72.3--230.8 MHz range.

Fig.~\ref{fig:rfi-per-band} shows the overal detected RFI occupancy per sub-band, as calculated over all observations except GLEAM 2014-03-17. The latter is the only observation affected by interference from digital video broadcasting (DVB), and we will analyse that observation later. The FM bands around 100~MHz and the ORBCOMM bands around 138~MHz are clearly present in the data. Excluding the RFI from DVB, the EoR high band is again slightly cleaner than the EoR low band, as its worst subband at 188.2 MHz has 1.03\% occupancy. The sub-bands at 145.9 and 149.8~MHz in the EoR low band have both 2.1\% occupancy. TODO what are these transmitters?

@@

Although the flagger does not adequately flag DVB, the right plot of Fig.~\ref{fig:dvb} shows that it is easy to recognize DVB interference from the data statistics. An option is to test whether the visibility RMS of a snapshot in any of the occupied DVB bands exceeds the RMS in DVB band 4, which is not used for broadcasting. When DVB is detected in a snapshot, in most cases the best action will be to remove the entire snapshot, as it is not worth having special mechanisms in place to salvage the small amount of residual good data within an affected snapshot. Because 2 out of 7 hours are affected in 1 out of the 9 randomly selected nights, the probability of observing DVB is roughly 3\%. This is detailed in the last column of Table~\ref{tbl:obs-list}.

@@

\subsection{Computational performance of \textsc{cotter}}
Because \textsc{cotter} processes the data at the highest resolution, its computational performance is an important consideration. Because of the large volume of MWA data, its runtime is strongly dependent on the input-output (IO) performance of the host system. Excluding IO, the main computational performance of \textsc{cotter} consists of running \textsc{aoflagger} on the data and collecting the statistics; performing its other tasks such as applying the cable delays and averaging increases the runtime by approximately a percent. In \citet{lofar-radio-environment} it is shown that the performance of the \textsc{aoflagger} is in normal situations dominated by IO, and hence the same holds for \textsc{cotter}: on a high-end desktop with 32 GB of memory and a 3.20-GHz Intel Core i7-3930K processor with six cores, and with a 5 disc RAID setup, processing a single 2~min snapshot with a 0.5~s / 40~kHz input resolution, using common averaging settings to output at 2~s / 80~kHz resolution, approximately 3 min are spend on reading the data, 5 min on RFI detection and 3.5 min on writing the data. Real-time processing can therefore be achieved by using 6 of such nodes in parallel.

\begin{itemize}
 \item Describe RFI examples (EoR low band \& high band)
 \item Performance
 \item Log N log S Histograms
\end{itemize}

>>>>

\begin{table*}
\caption{List of observations used in the analyses. The RFI column contains the fraction of visibilities that the initial \textsc{aoflagger} run has classified as RFI (which includes false positives). The DVB column displays the fraction of visibilities that were unusable because of the presence of DVB signals, and which the flagger does not flag. Not all observations cover the DVB frequencies. }\label{tbl:obs-list}
\begin{center}%
\begin{tabular}{|l|l|r|r|r|c|}
\hline
\textbf{Project} &\textbf{Date} & \textbf{Frequencies (MHz)} & \textbf{Duration} & RFI & DVB \\
\hline
EoR high& 2013-08-23 & 167.0--197.7              & 3h & 0.53\% & 0\% \\
GLEAM & 2013-08-25 & 72.3--133.8, 138.9--230.8 & 7h & 0.94\% & 0\% \\
EoR low& 2013-08-26 & 138.9--169.6              & 6h & 0.81\% & ---\\
GLEAM & 2013-11-05 & 72.3--133.8, 138.9--230.8 & 7h & 1.27\% & 0\% \\
GLEAM & 2013-11-25 & 72.3--133.8, 138.9--230.8 & 7h & 0.69\% & 0\% \\
EoR high& 2014-02-05 & 167.0--197.7              & 6h & 0.54\% & 0\%\\
GLEAM & 2014-03-16 & 72.3--133.8, 138.9--230.8 & 7h & 0.79\% & 0\% \\
GLEAM & 2014-03-17 & 72.3--133.8, 138.9--230.8 & 7h & 1.64\% & 30\% \\
EoR high& 2014-04-10 & 167.0--197.7              & 6h & 0.68\% & 0\%\\
GLEAM & 2014-06-18 & 72.3--133.8, 138.9--230.8 & 7h & 0.98\% & 0\% \\
Orbcomm RFI test&2014-08-27& 131.2--161.9     & 2$\times$8 m& 1.85\% & --- \\
\hline
\multicolumn{3}{|r|}{\textbf{Total}} & \textbf{63h} & \textbf{0.96\%} & \textbf{3\%}\\
\hline
\multicolumn{4}{|r|}{\textbf{With uniform channel coverage}} & \textbf{1.13\%} & \\
\hline
\end{tabular}
\end{center}
\end{table*}

@@

The fifth column of Table~\ref{tbl:obs-list} lists for each of the observation nights the ratio of samples that were classified as RFI by the \textsc{aoflagger}. This includes false detections, which are estimated in \citet{lofar-radio-environment} to account for approximately 0.4\% detections. Compared to the EoR observations, the GLEAM observations have a higher average of 1.05\% of RFI. This is mostly caused by the FM bands, which are only observed in GLEAM observations. The EoR low-band night shows 0.81\% RFI, while the EoR high-band observations have an average of 0.58\%. The total RFI occupancy in all observations is 0.96\%. Weighting the occupancy in each channel by the inverse time that it is observed results in an occupancy of 1.13\%, which is the global expected RFI ratio in the 72.3--230.8 MHz range.

Fig.~\ref{fig:rfi-per-band} shows the overal detected RFI occupancy per sub-band, as calculated over all observations except GLEAM 2014-03-17. The latter is the only night affected by interference from digital video broadcasting (DVB), and will be analysed later. The FM bands around 100~MHz and the ORBCOMM bands around 138~MHz are clearly present in the data. Excluding the RFI from DVB, the EoR high band is again slightly cleaner than the EoR low band, as its worst subband at 188.2 MHz has 1.03\% occupancy. The sub-bands at 145.9 and 149.8~MHz in the EoR low band have both 2.1\% occupancy. TODO what are these transmitters?

@@

Although the flagger does not adequately flag DVB, the right plot of Fig.~\ref{fig:dvb} shows that it is possible to determine whether a snapshot is affected by DVB interference from the data statistics. An option is to test whether the visibility RMS of a snapshot in any of the occupied DVB bands exceeds the RMS in DVB band 4, which is not used for broadcasting. These statistics are collected by \textsc{cotter}. When DVB is detected in a snapshot, the entire snapshot can be removed or the residual unaffected data within the snapshot can be used. However, it might not be worth it to have special mechanisms, such as having to account for lower SNR and a decreased $uv$ coverage, in place to salvage the residual good data within an affected snapshot. Because 2 out of 7 hours are affected in 1 out of the 9 randomly selected nights, the probability of observing DVB is roughly 3\%. This is detailed in the last column of Table~\ref{tbl:obs-list}.

@@

\subsection{Computational performance of \textsc{cotter}}
Because \textsc{cotter} processes the data at high resolution, its computational performance is an important consideration. A major contribution to the runtime is the reading and writing of the data, and the runtime is thus influenced by the input-output (IO) performance of the host system. Excluding IO, the main computational performance of \textsc{cotter} consists of running \textsc{aoflagger} on the data and collecting the statistics; performing all its other tasks such as applying the cable delays and averaging increases in total the runtime by approximately a percent. On a high-end desktop with 32 GB of memory and a 3.20-GHz Intel Core i7-3930K processor with six cores, and with a 5 disc RAID setup, processing a single 2~min snapshot with a 0.5~s / 40~kHz input resolution, using common averaging settings to output at 2~s / 80~kHz resolution, approximately 3~min are spend on reading the data, 5~min on RFI detection and 3.5~min on writing the data. Real-time processing can therefore be achieved by using 6 of such nodes in parallel.

\begin{itemize}
 \item Describe RFI examples (EoR low band \& high band)
 \item Log N log S Histograms
\end{itemize}
